GPU-Beschleunigung der Material Point Method

### 2018-11-23 mscthesis.tex
implicit midpoint rule, little problematic

Multi-component types are generally printed in bold letters $\boldsymbol{A}$, $\boldsymbol{a}$. Vectors furthermore use lower-case letters $\boldsymbol{v}$, while Matrices use upper-case letters $\boldsymbol{M}$. A variable in the Lagrangian form has a subscript 0 $ _0{\boldsymbol{x}}$, the Eulerian description will be written with an 'overline' $_t\boldsymbol{x}$ if distinction is needed. Occasionally notation may be omitted where it is apparent from context. Beginning from chapter \ref{sec:mat_point} variables defined on a particle will have subscript $p$ (e.g. $\boldsymbol{x}_p$). Grid cells will be assigned a subscript $i$ denoting the grid index (e.g. $\boldsymbol{x}_i$). This is not to be confused with the following Einstein-Notation. Therefore beginning with chapter \ref{sec:mat_point} Einstein-Notation will use greek letters ($\alpha,\beta$). A variable of the $n$-th time step with associated time $t^n = \sum_{i=1}^{n} t^i$ will have a right superscript: $\boldsymbol{x}^n$.

@@

m_i,

@@

  \frac{1}{\Delta t} ({m\boldsymbol{v}_i^{n+1}}-{m\boldsymbol{v}_i^{n}}) =

@@

  f_{i\alpha} \stackrel{\text{\ref{eq:piola},\ref{eq:volume_discr}}}{=}
  - \sum_p \frac{1}{J^n_p}P^n_{p\alpha\gamma}F^n_{p\beta\gamma}\frac{\partial w_{ip}}{\partial x_\beta}V^0_pJ^n_p

@@

  = - \sum_p P^n_{p\alpha\gamma}F^n_{p\beta\gamma}\frac{\partial w_{ip}}{\partial x_\beta}V^0_p

@@

>>>>

Multi-component types are generally printed in bold letters $\boldsymbol{A}$, $\boldsymbol{a}$. Vectors furthermore use lower-case letters $\boldsymbol{v}$, while Matrices use upper-case letters $\boldsymbol{M}$. A variable in the Lagrangian form has a subscript 0 $ _0{\boldsymbol{x}}$, the Eulerian description will be written with an 'overline' $_t\boldsymbol{x}$ if distinction is needed. Occasionally notation may be omitted where it is apparent from context. Beginning from chapter \ref{sec:mat_point} variables defined on a particle will have subscript $p$ (e.g. $\boldsymbol{x}_p$). Grid cells will be assigned a subscript $i$ denoting the grid index (e.g. $\boldsymbol{x}_i$). This is not to be confused with the following Einstein-Notation. Therefore beginning with chapter \ref{sec:mat_point} Einstein-Notation for components will use greek letters ($\alpha,\beta$). Occasionally summation over grid index $i$ and particles $p$ may be implied. A variable of the $n$-th time step with associated time $t^n = \sum_{i=1}^{n} t^i$ will have a right superscript: $\boldsymbol{x}^n$.

@@

m_i^n,

@@

  \frac{1}{\Delta t} (({m^n\boldsymbol{v}^{n+1}})_i-({m^n\boldsymbol{v}^n})_i) =

@@

\boldsymbol{f}^{n+\lambda}_i \stackrel{\text{\ref{eq:piola},\ref{eq:volume_discr}}}{=}
- \sum_p \frac{1}{J^n_p}\boldsymbol{P}^{n+\lambda}_{p}(\boldsymbol{F}^{n}_{p})^T \nabla w_{ip}V^0_pJ^n_p

@@

  = - \sum_p \boldsymbol{P}^{n+\lambda}_{p}(\boldsymbol{F}^{n}_{p})^T \nabla w_{ip}V^0_p.

@@

In eq. \ref{eq:evol_def_grad} the evolution of the deformation gradient is shown to be:
  \frac{\partial ^t_0\boldsymbol{F}}{\partial t} =  \nabla{_0\boldsymbol{v}}(_0\boldsymbol{x},t).
Discretizing the Lagrangian deformation gradient in time with eq. \ref{eq:midpoint} results in:
  \frac{\boldsymbol{F}^{n+1}_p + \boldsymbol{F}^{n}_p}{\Delta t} =  \nabla_0\boldsymbol{v}^{n+\lambda}(_0\boldsymbol{x}).
Pushing the right side forward to Eulerian view
  \frac{\boldsymbol{F}^{n+1}_p + \boldsymbol{F}^{n}_p}{\Delta t} =  \nabla_t\boldsymbol{v}^{n+\lambda}(_t\boldsymbol{x})^t_0\boldsymbol{F} = \nabla_t\boldsymbol{v}^{n+\lambda}(_t\boldsymbol{x})\boldsymbol{F}^n_p
and further applying the Galerkin discretization
  (_tv^{n+\lambda})_\alpha = (v_{i}^{n+\lambda})_\alpha w_i \Rightarrow \frac{\partial (_tv^{n+\lambda})_\alpha}{\partial x_\beta} =(v_i^{n+\lambda})_\alpha\frac{\partial w_i}{\partial x_\beta}
leads to the final update rule for the deformation gradient:
\boldsymbol{F}^{n+1}_p = \left( \boldsymbol{I} + \Delta t\sum_i \boldsymbol{v}_i^{n+\lambda}({\nabla w_{ip}})^T \right)\boldsymbol{F}^{n}_p.
\subsubsection{Grid nodal forces}
Starting from the last update on the grid which is typically the position with eq. \ref{eq:midpoint}:
\frac{\partial _t\boldsymbol{x}}{\partial t} = {_t\boldsymbol{v}} \Rightarrow \frac{\boldsymbol{\hat{x}}_i^{n+1} - \boldsymbol{x}_i^{n}}{\Delta t} = \boldsymbol{v}_i^{n+\lambda} =  (1-\lambda)\boldsymbol{v}_i^n + \lambda\boldsymbol{\hat{v}}_i^{n+1}
  \Leftrightarrow \boldsymbol{\hat{x}}_i^{n+1}   =\boldsymbol{x}_i^{n} + \Delta t \left((1-\lambda)\boldsymbol{v}_i^n + \lambda\boldsymbol{\hat{v}}_i^{n+1}\right).
Due to \ref{eq:space_discr} a velocity update can be put together as
  \frac{1}{\Delta t} (({m^n\boldsymbol{\hat{v}}^{n+1}})_i-({m^n\boldsymbol{v}^n})_i)
  = \boldsymbol{f}_i^{n+\lambda}
 \Leftrightarrow
 \boldsymbol{\hat{v}}^{n+1}_i
 = \boldsymbol{v}^n_i+ \frac{\Delta t}{m_i^n}\boldsymbol{f}_i^{n+\lambda},
where the force can be also described in terms of the in-between point:
 \boldsymbol{f}_i^{n+\lambda} = \boldsymbol{\sigma}^{n+\lambda}_p(\boldsymbol{F}_p^{n+\lambda}(\boldsymbol{\hat{x}}_i^{n+1})).
